### 2021-11-09 elsevier/appendix.Rmd
for Elsevier journals

---
output: pdf_document
---
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}

Fig.\ref{fig:s2021} is in Supplementary materials.
```{r s2021, echo=FALSE, fig.cap='A plot in Supplementary Materials', out.width="80%", fig.align='center'}
knitr::include_graphics(here::here("results", "pictures", "2021.png"))
```

Table.\ref{tab:stable} is in Supplementary materials.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
          group_by(supp) %>%
          summarise(median = median(len),
                    mean = mean(len)) %>% 
	kable("simple",
	      booktabs = T,
	      col.names = c("Supplement type", 
	      	    	"Median",
	      	    	"Mean"), 
	      align = "cll",
	      caption = "\\label{tab:stable}This table again")
```

Some random SAS code  
```
PROC MCMC 
	data=Data outpost=Dataoutput 
		nbi=1000000 
		nmc=1000000
		thin=10
		seed=1
		diag=all
		monitor=(p1 p2 cp I w); 
```