### 2021-11-09 elsevier/manuscript.Rmd

---
title: Article Name
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.  
  1. what is the study about?  
  2. what problem does it address?  
  3. how did you conduct the research?  
  4. what were the main findings?  
  5. why is it important?

keywords: "keywordA, keywordB"
journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
#linenumbers: true
numbersections: true
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage{lineno}\linenumbers
csl: harmful-algae.csl
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    includes:  
      in_header: my_header.tex
linkcolor: blue
---
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(here)
library(kableExtra)
library(tidyverse)

data <- read.csv(here("data", "process", "newdata.csv"))
```

# Introduction {#intro}
1. general information
1. research gap
1. research aim

# Materials and methods {#method}
## Adding figures {#figure}
Add Fig. \ref{fig:boxplot} here.

```{r boxplot, echo=FALSE, fig.cap='A boxplot', out.height="50%", fig.align='center'}
knitr::include_graphics(here::here("results", "figures", "boxplot.pdf"))
```

If the caption is too long (fig.\ref{fig:2021}), use the [text-reference](https://bookdown.org/yihui/rmarkdown/bookdown-markdown.html#text-references). 
See the code in "elsevier/manuscript.Rmd" for more information.

(ref:longcaption) This is a very long caption
```{r 2021, echo=FALSE, fig.cap='(ref:longcaption)', out.width="50%", fig.align='center'}
knitr::include_graphics(here::here("results", "pictures", "2021.png"))
```

## Adding tables {#table}
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
          group_by(supp) %>%
          summarise(median = median(len),
                    mean = mean(len)) %>% 
	kable("simple",
	      booktabs = T,
	      col.names = c("Supplement type", 
	      	    	"Median",
	      	    	"Mean"), 
	      align = "cll",
	      caption = "\\label{tab:cooltable}Mean and Median")
```

## Adding equation {#equation}
1. The _variable_ \(x\) and the __function__ \(f(x)\)
1. The *variable* $x$ and the **function** $f(x)$
1. superscript^2^
1. NO~2~, NO~3~, PO~4~, NH~4~

\[Y|X \sim Bernoulli(p)\]

or

\begin{equation}
\label{eq:cutoff}
p(x) = P(Y = 1|X = x) = 
\left\{
    \begin{array}{lr}
          p_1 = P(Y = 1|X \le cp), & \text{if } x \le cp\\
          p_2 = P(Y = 1|X > cp), & \text{if } x > cp
    \end{array}
\right.
\end{equation}

## Cross-reference
* figure: `\ref{fig:label}`
* table: `\ref{tab:label}`
* equation: `\ref{eq:label}`
* section: `\ref{label}`

**Note:** only alphanumeric characters (`a-z`, `A-Z`, `0-9`), `-`, `/`, and `:` are allowed in labels.

1. Fig.\ref{fig:boxplot} and fig.\ref{fig:2021}
1. Table. \ref{tab:cooltable}
1. Equation \ref{eq:cutoff}
1. Section \ref{intro} and section \ref{figure}

## Citation syntax
- cite directly @Davis2009
- put citations in parentheses [@Walls2018]
- cite multiple entries [@Davis2009; @Walls2018]
- suppress the mention of the author [-@Liu2011a]

# Results and discussion {#result}
## Use code inline
The maximum tooth length is `r max(data$len)`.

## Random things
- download `.csl` file at [Zotero Style Repository](https://www.zotero.org/styles)
- *References* section is created at the end of the document by default. 
To put *References* section in a specific place (e.g. before Supplementary Materials):

```
# References
<div id="refs"></div>
# Supplementary Materials
```
- check spelling in rmarkdown: `F7`
- [word count addin](https://github.com/benmarwick/wordcountaddin)

# Conclusions {#conclusion}

# CRediT authorship contribution statement {#authorship .unnumbered}

# Data availability {#datasharing .unnumbered}

# Declaration of competing interest {#declaration .unnumbered}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

# Acknowledgements {#thanks .unnumbered}

# Supplementary materials {#appendix .unnumbered}
```{r, child="appendix.Rmd"}

```

# References {#references .unnumbered}