### 2021-11-09 elsevier/manuscript.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Article Name},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing

% Pandoc header
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Article Name}
    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
   \ead{alice@example.com} 
    \author[Another University]{Bob Security}
   \ead{bob@example.com} 
    \author[Another University]{Cat Memes\corref{2}}
   \ead{cat@example.com} 
    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
   \ead{derek@example.com} 
      \address[Some Institute of Technology]{Department, Street, City, State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  This is the abstract.\\
  1. what is the study about?\\
  2. what problem does it address?\\
  3. how did you conduct the research?\\
  4. what were the main findings?\\
  5. why is it important?
  \end{abstract}
   \begin{keyword} keywordA, keywordB\end{keyword}
 \end{frontmatter}

\section{Introduction}\label{intro}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  general information
\item
  research gap
\item
  research aim
\end{enumerate}

\section{Materials and methods}\label{method}

\subsection{Adding figures}\label{figure}

Add Fig. \ref{fig:boxplot} here.

\begin{figure}

{\centering \includegraphics[height=0.5\textheight]{C:/Users/LY/Google Drive/Git/writing_journal_article_in_rmarkdown/results/figures/boxplot}

}

\caption{A boxplot}\label{fig:boxplot}
\end{figure}

If the caption is too long (fig.\ref{fig:2021}), use the
\href{https://bookdown.org/yihui/rmarkdown/bookdown-markdown.html\#text-references}{text-reference}.
See the code in ``elsevier/manuscript.Rmd'' for more information.

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{C:/Users/LY/Google Drive/Git/writing_journal_article_in_rmarkdown/results/pictures/2021}

}

\caption{This is a very long caption}\label{fig:2021}
\end{figure}

\subsection{Adding tables}\label{table}

\begin{longtable}[]{@{}cll@{}}
\caption{\label{tab:unnamed-chunk-2}\label{tab:cooltable}Mean and
Median}\tabularnewline
\toprule
Supplement type & Median & Mean\tabularnewline
\midrule
\endfirsthead
\toprule
Supplement type & Median & Mean\tabularnewline
\midrule
\endhead
OJ & 22.7 & 20.66333\tabularnewline
VC & 16.5 & 16.96333\tabularnewline
\bottomrule
\end{longtable}

\subsection{Adding equation}\label{equation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The \emph{variable} \(x\) and the \textbf{function} \(f(x)\)
\item
  The \emph{variable} \(x\) and the \textbf{function} \(f(x)\)
\item
  superscript\textsuperscript{2}
\item
  NO\textsubscript{2}, NO\textsubscript{3}, PO\textsubscript{4},
  NH\textsubscript{4}
\end{enumerate}

\[Y|X \sim Bernoulli(p)\]

or

\begin{equation}
\label{eq:cutoff}
p(x) = P(Y = 1|X = x) = 
\left\{
    \begin{array}{lr}
          p_1 = P(Y = 1|X \le cp), & \text{if } x \le cp\\
          p_2 = P(Y = 1|X > cp), & \text{if } x > cp
    \end{array}
\right.
\end{equation}

\subsection{Cross-reference}\label{cross-reference}

\begin{itemize}
\tightlist
\item
  figure: \texttt{\textbackslash{}ref\{fig:label\}}
\item
  table: \texttt{\textbackslash{}ref\{tab:label\}}
\item
  equation: \texttt{\textbackslash{}ref\{eq:label\}}
\item
  section: \texttt{\textbackslash{}ref\{label\}}
\end{itemize}

\textbf{Note:} only alphanumeric characters (\texttt{a-z}, \texttt{A-Z},
\texttt{0-9}), \texttt{-}, \texttt{/}, and \texttt{:} are allowed in
labels.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Fig.\ref{fig:boxplot} and fig.\ref{fig:2021}
\item
  Table. \ref{tab:cooltable}
\item
  Equation \ref{eq:cutoff}
\item
  Section \ref{intro} and section \ref{figure}
\end{enumerate}

\subsection{Citation syntax}\label{citation-syntax}

\begin{itemize}
\tightlist
\item
  cite directly Davis et al. (2009)
\item
  put citations in parentheses (Walls et al., 2018)
\item
  cite multiple entries (Davis et al., 2009; Walls et al., 2018)
\item
  suppress the mention of the author (2011)
\end{itemize}

\section{Results and discussion}\label{result}

\subsection{Use code inline}\label{use-code-inline}

The maximum tooth length is 33.9.

\subsection{Random things}\label{random-things}

\begin{itemize}
\tightlist
\item
  download \texttt{.csl} file at
  \href{https://www.zotero.org/styles}{Zotero Style Repository}
\item
  \emph{References} section is created at the end of the document by
  default. To put \emph{References} section in a specific place
  (e.g.~before Supplementary Materials):
\end{itemize}

\begin{verbatim}
# References
<div id="refs"></div>
# Supplementary Materials
\end{verbatim}

\begin{itemize}
\tightlist
\item
  check spelling in rmarkdown: \texttt{F7}
\item
  \href{https://github.com/benmarwick/wordcountaddin}{word count addin}
\end{itemize}

\section{Conclusions}\label{conclusion}

\section*{CRediT authorship contribution statement}\label{authorship}
\addcontentsline{toc}{section}{CRediT authorship contribution statement}

\section*{Data availability}\label{datasharing}
\addcontentsline{toc}{section}{Data availability}

\section*{Declaration of competing interest}\label{declaration}
\addcontentsline{toc}{section}{Declaration of competing interest}

The authors declare that they have no known competing financial
interests or personal relationships that could have appeared to
influence the work reported in this paper.

\section*{Acknowledgements}\label{thanks}
\addcontentsline{toc}{section}{Acknowledgements}

\section*{Supplementary materials}\label{appendix}
\addcontentsline{toc}{section}{Supplementary materials}

\renewcommand{\thefigure}{S\arabic{figure}}

\setcounter{figure}{0} \renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0} \renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}

Fig.\ref{fig:s2021} is in Supplementary materials.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{C:/Users/LY/Google Drive/Git/writing_journal_article_in_rmarkdown/results/pictures/2021}

}

\caption{A plot in Supplementary Materials}\label{fig:s2021}
\end{figure}

Table.\ref{tab:stable} is in Supplementary materials.

\begin{longtable}[]{@{}cll@{}}
\caption{\label{tab:unnamed-chunk-4}\label{tab:stable}This table
again}\tabularnewline
\toprule
Supplement type & Median & Mean\tabularnewline
\midrule
\endfirsthead
\toprule
Supplement type & Median & Mean\tabularnewline
\midrule
\endhead
OJ & 22.7 & 20.66333\tabularnewline
VC & 16.5 & 16.96333\tabularnewline
\bottomrule
\end{longtable}

Some random SAS code

\begin{verbatim}
PROC MCMC 
    data=Data outpost=Dataoutput 
        nbi=1000000 
        nmc=1000000
        thin=10
        seed=1
        diag=all
        monitor=(p1 p2 cp I w); 
\end{verbatim}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Davis2009}{}
Davis, T.W., Berry, D.L., Boyer, G.L., Gobler, C.J., 2009. The effects
of temperature and nutrients on the growth and dynamics of toxic and
non-toxic strains of Microcystis during cyanobacteria blooms. Harmful
Algae 8, 715--725. \url{https://doi.org/10.1016/j.hal.2009.02.004}

\hypertarget{ref-Liu2011a}{}
Liu, X., Lu, X., Chen, Y., 2011. The effects of temperature and nutrient
ratios on Microcystis blooms in Lake Taihu, China: An 11-year
investigation. Harmful Algae 10, 337--343.
\url{https://doi.org/10.1016/j.hal.2010.12.002}

\hypertarget{ref-Walls2018}{}
Walls, J.T., Wyatt, K.H., Doll, J.C., Rubenstein, E.M., Rober, A.R.,
2018. Hot and toxic: Temperature regulates microcystin release from
cyanobacteria. Science of the Total Environment 610-611, 786--795.
\url{https://doi.org/10.1016/j.scitotenv.2017.08.149}

\end{document}