### 2021-11-24 elsevier/cover_letter.Rmd

---
output: 
  pdf_document: default
  bookdown::word_document2: default  # for cross-references
---
# Cover letter {-}

### New submission {-}
Month Day, Year

Dear Dr. AAA,

I am happy to submit my manuscript, **article_name**, for your consideration at *journal_name*. 
This work *do sth interesting*. 
The main conclusion is that *sth cool*.

All of the authors have read and approved the paper and it has not been published previously nor is it being considered by any other peer-reviewed journal.

The manuscript has also been submitted to bioRxiv as a preprint.

Sincerely,

Author_name, PhD  
Professor

### Resubmissions {-}
Month Day, Year

Dear Dr. BBB,

I am happy to resubmit my manuscript, **article_name**, for your reconsideration at *journal_name*. 
I am grateful to you and the reviewers who were very encouraging about the content of the manuscript.  
I apologize for taking so long to resubmit.  
Too many things got in the way over the past few months.

All of the authors have read and approved the paper and it has not been published previously nor is it being considered by any other peer-reviewed journal.

The manuscript was previously submitted to bioRxiv as a preprint.

Sincerely,

Author_name, PhD  
Professor