### 2021-11-24 elsevier/manuscript.Rmd

---
title: Article Name
author:
  - name: Author One
    email: one@example.com
    affiliation: a,1
  - name: Author Two
    email: two@example.com
    affiliation: a,b
    footnote: 1
  - name: Author Three
    email: three@example.com
    affiliation: c
    footnote: 2
  - name: Author Four
    email: four@example.com
    affiliation: a
    footnote: 2
address:
  - code: a
    address: Graduate School of Engineering, Nagasaki University, 1-14 Bunkyo-machi, Nagasaki 852-8521, Japan
  - code: b
    address: Department, University, Street, City, State, Zip
  - code: c
    address: Department, University, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
  - text: "^1^Present address: Department, University, Street, City, State, Zip"
abstract: |
  This is the abstract.  
  1. what is the study about?  
  2. what problem does it address?  
  3. how did you conduct the research?  
  4. what were the main findings?  
  5. why is it important?

keywords: "keywordA, keywordB"
journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
#linenumbers: true
numbersections: true
layout: review      #review=doublespace, preprint=single, 5p=2column
csl: harmful-algae.csl        #download .csl at www.zotero.org/styles
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    includes:  
      in_header: my_header.tex
  bookdown::word_document2: default  # for cross-references
linkcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(kableExtra)
```

```{r}
data <- read.csv(here("data", "process", "newdata.csv"))
```

\newpage

# Introduction {#intro}
1. general information
1. research gap
1. research aim

# Materials and methods {#method}
## Adding figures {#figure}
Add Fig. \ref{fig:boxplot} here.

```{r boxplot, fig.cap='A boxplot', out.height="50%", fig.align='center'}
knitr::include_graphics(here::here("results", "figures", "boxplot.pdf"))
```

If the caption is too long (fig.\ref{fig:2021}), use the [text-reference](https://bookdown.org/yihui/rmarkdown/bookdown-markdown.html#text-references). 
See the code in "elsevier/manuscript.Rmd" for more information.

(ref:longcaption) This is a very long caption
```{r 2021, fig.cap='(ref:longcaption)', out.width="50%", fig.align='center'}
knitr::include_graphics(here::here("results", "pictures", "2021.png"))
```

## Adding tables {#table}
```{r}
data %>% 
          group_by(supp) %>%
          summarise(median = median(len),
                    mean = mean(len)) %>% 
	kable("simple",
	      booktabs = T,
	      col.names = c("Supplement type", 
	      	    	"Median",
	      	    	"Mean"), 
	      align = "cll",
	      caption = "\\label{tab:cooltable}Mean and Median")
```

## Adding equation {#equation}
1. The _variable_ \(x\) and the __function__ \(f(x)\)
1. The *variable* $x$ and the **function** $f(x)$
1. superscript^2^
1. NO~2~, NO~3~, PO~4~, NH~4~

\[Y|X \sim Bernoulli(p)\]

or

\begin{equation}
\label{eq:cutoff}
p(x) = P(Y = 1|X = x) = 
\left\{
    \begin{array}{lr}
          p_1 = P(Y = 1|X \le cp), & \text{if } x \le cp\\
          p_2 = P(Y = 1|X > cp), & \text{if } x > cp
    \end{array}
\right.
\end{equation}

## Cross-reference
* figure: `\ref{fig:label}`
* table: `\ref{tab:label}`
* equation: `\ref{eq:label}`
* section: `\ref{label}`

**Note:** only alphanumeric characters (`a-z`, `A-Z`, `0-9`), `-`, `/`, and `:` are allowed in labels.

1. Fig.\ref{fig:boxplot} and fig.\ref{fig:2021}
1. Table. \ref{tab:cooltable}
1. Equation \ref{eq:cutoff}
1. Section \ref{intro} and section \ref{figure}

## Citation syntax
- cite directly @Davis2009
- put citations in parentheses [@Walls2018]
- cite multiple entries [@Davis2009; @Walls2018]
- suppress the mention of the author [-@Liu2011a]

# Results {#result}
## Use code inline
The maximum tooth length is `r max(data$len)`.

## Random things
- download `.csl` file at [Zotero Style Repository](https://www.zotero.org/styles)
- *References* section is created at the end of the document by default. 
To put *References* section in a specific place (e.g. before Supplementary Materials):

```
# References
<div id="refs"></div>
# Supplementary Materials
```
- check spelling in rmarkdown: `F7`
- [word count addin](https://github.com/benmarwick/wordcountaddin)

# Discussion {#discussion}

# Conclusions {#conclusion}

\newpage

```{r, child="interest.Rmd"}

```

```{r, child="credit.Rmd"}

```

```{r, child="thanks.Rmd"}

```

\newpage
# References {#references .unnumbered}
<div id="refs"></div>

\newpage
```{r, child="appendix.Rmd"}

```

\newpage
```{r, child="highlight.Rmd"}

```

# Graphical abstract {#graphic .unnumbered}
Delete `eval=FALSE` before run the code chunk

```{r, out.width="50%", fig.align='center', eval=FALSE}
knitr::include_graphics(here("results", "figures", "graphical_abstract.pdf"))
```

\newpage
```{r, child="cover_letter.Rmd"}

```

```{r, child="response_to_reviewers.Rmd"}

```