### 2022-02-11 elsevier/appendix.Rmd
Formatting Appendix

---
output: 
  bookdown::pdf_book: default
  bookdown::word_document2: default  # for cross-references
---
```{r setappendix, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(kableExtra)
```

# (APPENDIX) Appendix {-}

\renewcommand{\thefigure}{A.\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{A.\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{A.\arabic{equation}}
\setcounter{equation}{0}

# Supplementary materials A {#appendixA}
## A cool figure

Fig.\ref{fig:s2021} is in Supplementary materials.

```{r s2021, echo=FALSE, fig.cap='A plot in Supplementary Materials', out.width="80%", fig.align='center'}
knitr::include_graphics(here::here("results", "pictures", "2021.png"))
```

## An awesome table
Table.\ref{tab:stable} is in Supplementary materials.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
data %>% 
          group_by(supp) %>%
          summarise(median = round(median(len), 3),
                    mean = round(mean(len), 3)) %>% 
	kable("simple",
	      booktabs = T,
	      col.names = c("Supplement type", 
	      	    	"Median",
	      	    	"Mean"), 
	      align = "cll",
	      caption = "\\label{tab:stable}This table again")
```

\newpage

\renewcommand{\thefigure}{B.\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{B.\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{B.\arabic{equation}}
\setcounter{equation}{0}

# Supplementary materials B {#appendixB}
## Some random code

Some random SAS code  
```
PROC MCMC 
	data=Data outpost=Dataoutput 
		nbi=1000000 
		nmc=1000000
		thin=10
		seed=1
		diag=all
		monitor=(p1 p2 cp I w); 
```

## A green photo

```{r green, echo=FALSE, fig.cap='A green photo', out.width="80%", fig.align='center'}
knitr::include_graphics(here::here("results", "pictures", "green.jpeg"))
```