### 2024-05-18 elsevier/cover_letter.Rmd
update `cover_letter.Rmd`: add long version for new submission

---
output: 
  pdf_document: default
  bookdown::word_document2: default  # for cross-references
---
# Cover letter {-}

### New submission - short version {-}
Month Day, Year

Dear Dr. AAA,

I am happy to submit my manuscript, **article_name**, for your consideration at *journal_name*. 
This work *did sth interesting*. 
The main conclusion is that *sth cool*.

All of the authors have read and approved the paper and it has not been published previously nor is it being considered by any other peer-reviewed journal.

The manuscript has also been submitted to bioRxiv as a preprint.

Sincerely,

Author_name, PhD  
Professor

### New submission - long version {-}
Journal editor's name  
Title  
*Journal Name*

Month Day, Year

Dear Dr.[last_name],

I would like to submit the manuscript entitled [**article_name**] by [author names] for consideration as [paper type: an original article / a brief communication / a case report / etc.] in [*Journal Name*].

[Summarize the research problem/gap, your main research findings, and the implications of your findings]. 
We believe these findings will be of interest to the readers of your journal.

[[Revise or delete any of the following standard statements used in cover letter]]  
We declare that this manuscript is original, has not been published before, and is not currently under consideration for publication elsewhere.

We are unaware of any conflicts of interest associated with this publication, and there has been no significant financial support for this work that could have influenced its outcome. 
As the Corresponding Author, I confirm that the manuscript has been read and approved for submission by all the named authors.

[[Add any other statements/information required by the journal here]]

We hope you find our manuscript suitable for publication and look forward to your response.

Sincerely,

[Corresponding Author's Name], PhD  
Department Name, Institution Name  
Postal Address  
Tel./Fax: xxxxxx  
Email: xxxxxx

### Resubmissions {-}
Month Day, Year

Dear Dr. BBB,

I am happy to resubmit my manuscript, **article_name**, for your reconsideration at *journal_name*. 
I am grateful to you and the reviewers who were very encouraging about the content of the manuscript.  
I apologize for taking so long to resubmit.  
Too many things got in the way over the past few months.

All of the authors have read and approved the paper and it has not been published previously nor is it being considered by any other peer-reviewed journal.

The manuscript was previously submitted to bioRxiv as a preprint.

Sincerely,

Author_name, PhD  
Professor